\documentclass[twocolumn]{article}

\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}

\newcommand{\fix}[1]{{\bf #1}}

\setcounter{secnumdepth}{2}

\title{Playing to Program:  An Intelligent Programming Tutor for RUR-PLE}
\author{Marie desJardins and an Ensemble of Students \\
University of Maryland, Baltimore County \\ 
1000 Hilltop Circle \\
Baltimore MD  21250 \\
mariedj@cs.umbc.edu
}


\begin{document}
\maketitle

\begin{abstract}
Intelligent tutoring systems (ITSs) have proven their effectiveness in
contributing to student learning.  ITSs are automated programs
that provide students with a one-on-one tutor, allowing them to work
at their own pace, so they can spend more time on their weaker areas
of the subject matter.  RUR--Python Learning Environment (RUR-PLE), a
virtual environment to help 
students learn to program in Python, provides an interface for
students to write their own Python code and then be presented with a
visualization of that same code [CITE].  
The RUR-PLE system provides a sequence of learning lessons for
students to explore.  We have extended RUR-PLE to provide an
intelligent tutoting system interface that consists of three
components:  (1) a
student model that tracks student understanding, (2) a diagnosis module
that provides tailored feedback to students, and (3) a problem selection
module that guides the student's learning process.  In this paper, we
describe the basis RUR-PLE system and our extensions, and present the
results of a user study in which we evaluated the effectiveness of our
three ITS modules.
\end{abstract}

\section{Introduction and Related Work}
\label{sec:intro}

\fix{ADD introduction and motivation.}

\section{Infrastructure}
\label{sec:student-model}

\fix{ADD description of concept map, how we built/tested it 
(i.e., justification for these concepts and why they're
connected as they are, and how we instantiate and track it for a
specific user (presumably using some kind of Bayesian updating).  
Idea:  validate/finalize it by some testing process on
a group of students (i.e., is it in fact the case that students in
general need to understand concept X before they can apply concept Y)
-- use a problem suite (where each problem includes known concepts)
to measure these dependencies.} 


\section{Pre-test and Problem Selection}
\label{sec:prob-selection}

\fix{ADD: How is the student model used to generate and select
  problems for the student to work on?  What is the motivation for our
  approach, and how does it work?}

\section{Experimental Design}
\label{sec:exper-design}

\fix{ADD:  Methodology:  design of the user study (set up in such a}

\section{Conclusions and Future Work}
\label{conclusions}

\fix{ADD:  What have we contributed?  What are the take-away lessons?
  What would we work on next?}
  
\section*{References}


\end{document}
