\documentclass[twocolumn]{article}

\usepackage{aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}

\newcommand{\fix}[1]{{\bf #1}}

\setcounter{secnumdepth}{2}

\title{Playing to Program:  An Intelligent Programming Tutor for RUR-PLE}
\author{Marie desJardins and an Ensemble of Students \\
University of Maryland, Baltimore County \\ 
1000 Hilltop Circle \\
Baltimore MD  21250 \\
mariedj@cs.umbc.edu
}


\begin{document}
\maketitle

\begin{abstract}
Intelligent tutoring systems (ITSs) have proven their effectiveness in
contributing to student learning.  ITSs are automated programs
that provide students with a one-on-one tutor, allowing them to work
at their own pace, so they can spend more time on their weaker areas
of the subject matter.  RUR--Python Learning Environment (RUR-PLE), a
virtual environment to help 
students learn to program in Python, provides an interface for
students to write their own Python code and then be presented with a
visualization of that same code [CITE].  
The RUR-PLE system provides a sequence of learning lessons for
students to explore.  We have extended RUR-PLE to provide an
intelligent tutoring system interface that consists of three
components: 
 (1) a student model that tracks student understanding, 
(2) a diagnosis module
that provides tailored feedback to students, and 
(3) a problem selection
module that guides the student's learning process.  In this paper, we
describe the basis RUR-PLE system and our extensions, and present the
results of a user study in which we evaluated the effectiveness of our
three ITS modules.
\end{abstract}

\section{Introduction and Related Work}
\label{sec:intro}

\fix{ADD introduction and motivation.}

\section{Infrastructure}
\label{sec:student-model}

The program infrastructure provides the ability for user generation of programs. Upon first starting the system, a series of forms are used to obtain information about a student's background, including course history and course grades. Once logged in, a tabular interface is presented. The start tab presents a general instruction sheet along with a list and description of basic RUR-PLE functions. On the following tab, a problem is presented to the user with detailed instructions on the behavior expected in the associated problems RUR-PLE World upon successful completion of the problem. Each time a test run or a final submission is performed, the users code is appended to a file along with the test run or submission time. The edited world file is also saved to align the test run with the environment changes to the RUR-PLE World. The final tab provides access to a full Python interpreter.


\section{Pre-test and Problem Selection}
\label{sec:prob-selection}

We designed a pretest to administer to students that will evaluate their understanding of various concepts.  This pretest consists of twenty multiple-choice questions each of which assesses student understanding across a variety of concepts.  The concepts we include in our evaluation are loops, function calls, conditionals and variables.  As the students answer various questions correctly or incorrectly, we are able to predict their level of understanding of the various concepts from the accuracy of their responses.  We designed pretest problems at a variety of difficulty levels in order to detect fine grained levels of comprehension.

The problems we presented to the students were more complex, as they involved writing programs to solve problems.  These tasks also incorporated different concepts at different levels of difficulty.  

\section{Experimental Design}
\label{sec:exper-design}
 
Twenty-four participants were recruited for the study. Participants were UMBC students who had completed at least one computer science major class but have not taken any 300 level computer science classes.
The participants were asked to take a pretest (described above). The participant were presented with 10 programming problems to solve using a PUFFL. The participant  were able to skip a problem if they had trouble solving it.


\section{Conclusions and Future Work}
\label{conclusions}

\fix{ADD:  What have we contributed?  What are the take-away lessons?
  What would we work on next?}
  
\section*{References}


\end{document}
